Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-25 3dimpvt/_region_.tex
completed v2 changes
fixed first half, need more refs and waiting for second half
next draft

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\message{ !name(paper.tex) !offset(204) }
\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{For each tile, we want to select an image such that $d$ and $\alpha$ are minimized}
  \label{fig:scoringFunction}
\end{figure}

%\begin{figure}
%  \centering
%  \includegraphics[height=1.5in]{directMapping.png}
%  \caption{The result of direct texture mapping based on locally
%    optimized textures.}
%  \label{fig:directMapping}
%\end{figure}
\message{ !name(paper.tex) !offset(699) }

>>>>

% Pages are numbered in submission mode, and unnumbered in
% camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\message{ !name(paper.tex) !offset(509) }
\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a): Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b): The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c): Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}
\message{ !name(paper.tex) !offset(734) }

### 2012-06-25 3dimpvt/paper.tex
next draft

In this paper, we present a number of approaches to texture mapping 3D models of indoor environments made of planar surfaces in the presence of uncertainty and noise in camera poses. In particular, we consider a human-operated backpack system with a number of laser range scanners as well as 2 cameras facing left and right, each equipped with fisheye lenses reaching an approximately 180$^{\circ}$ field of view. Considering that the human operator walks at about 1 meter/second, and images are captured at 5 Hz, there is a considerable amount of overlap between successive images.

With the laser scanners active, the human operator wearing the
backpack takes great care to walk a path such that every wall in the
desired indoor environment is traversed and scanned lengthwise at
least once.

Using data collected by the onboard sensors and applying multiple localization
and loop-closure algorithms \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is then localized over its
data collection period. This requires recovering the 6 degrees of freedom for the backpack as well as all its sensors, including the cameras mounted on it. \footnote{In this paper, we use the terms localization and pose recovery interchangeably, in that they both refer to recovering position and orientation.} Once this is complete, the data from the laser range scanners is used to generate a 3D point cloud of the surrounding environment. 
Approximate normal vectors for each point in the
point cloud are then calculated by gathering neighboring points within
a small radius and processing them through principal component
analysis. These normal vectors allow for the classification and
grouping of adjacent points into structures such as walls, ceilings,
floors, and staircases. A RANSAC algorithm is then employed to fit
polygonal planes to these structured groupings of points, resulting in
a fully planar model \cite{sanchez2012point}. This model, consisting
of multiple 2D polygonal planes in 3D space, along with the set of
images captured by the backpack's cameras and their noisy 3D poses,
can be considered the input to our texture mapping problem.

@@

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the projection method above. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane, all in world coordinates
\cite{rayintersection}.

@@

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{We wish to minimize camera angle $\alpha$ and distance $d$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_naive.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_cache_full.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_cache_full_shifted.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_dynprog_noblend.jpg}}
  \caption{(a): Direct mapping. (b): Mapping with caching. (c):
    Mapping with caching after image alignment. (d): Seam minimization approach (including image alignment)}
  \label{fig:compareAll}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images chosen by its neighbors, using
similar images is likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function. Before reusing
this image, we ensure it meets a threshold, which we set
unconditionally to $\alpha < 45^\circ$, to be considered a viable
image, with $\alpha$ as the camera angle as shown in Figure
\ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we then check
our second subset of images, which consists of images that were taken
near the images in the first subset, both spatially and
temporally. These images are not the same as the ones used for
neighboring tiles, but they were taken at a similar location and time,
suggesting that their localization and projection are very
similar. Again, if no good image is found according to the same
threshold, we must then search the entire set of candidate images,
selecting based on the criteria from Figure \ref{fig:scoringFunction}.

@@

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching (Section
\ref{sec:directMapping}. This set of images constitutes a good
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on our planeis covered by
at least one image in this set; this ensures no holes in our final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring good camera
distances and angles. Third, as a side result of the scoring function,
selected images are only viable candidates for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards
refining and combining their projections.

@@

Figure \ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane, despite the fact that it was
generated after careful hand tuning. Many planes with fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

@@

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we perform image rotation and shifting
in order to maximize SIFT feature matches in overlapping areas between
images. We then project these images onto our plane, applying textures
either with the tile caching method from Section
\ref{sec:mappingWithCaching}, or with the more specialized method to
be described in Section \ref{sec:seamMinimization}. This choice
depends on the availability of head-on images for a given plane. We
then finish by applying linear alphar blending to smooth out any
remaining seams.

@@

\subsection{Image Pose Refinement}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images using
SIFT feature matching \cite{lowe1999object}. The SIFT matches
determine $dx$ and $dy$ distances between each pair of features for
two images on the plane, though these distances may not always be the
same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that our SIFT-based distances
are reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $dx_{i,j}$ and $dy_{i,j}$ distances between two
images $i$ and $j$. The RANSAC framework handles the consensus-building machinery,
and requires a fitting function and a distance function. For this
application, the fitting function simply finds the average distance
between matches in a pair of images. The distance function for a pair
of points is chosen to be the difference between those points' SIFT match distance
and the average distance computed by the fitting function; we use a 10
pixel outlier threshold. This means that a SIFT match is labeled as an
outlier if its horizontal or vertical distance is not within 10 pixels
of the average distance computed by the fitting function.

@@

The $\vec{\beta}$ that minimizes $||X \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches were found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. Figures \ref{fig:compareAll}(c) and
\ref{fig:compareAll}(d) compare the tile caching method against this
seam minimization method. For the latter, we arbitrarily choose one image for texturing where images overlap. Though both methods provide aligned texturing, the seam minimization approach results in fewer
visible discontinuities, since it directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, seam minimization
guarantees the best selection of images, while the sequential tile
caching method may select images early on that turn out to be poor
choices once subsequent tiles have been processed. The seam
minimization approach is also less intensive in terms of memory usage
and runtime, as it does not require discretizing planes or images.

@@

\begin{figure*}
  \centering
  \includegraphics[width=6in]{pipelineimages.pdf}
  \caption{Textures corresponding to each step in our texture mapping
    pipeline.}
  \label{fig:pipelineimages}
\end{figure*}

@@

After blending pixels linearly across overlapping regions using alpha
blending, texture mapping is complete. Figure
\ref{fig:pipelineimages} shows each incremental step of our texture
mapping approach for a different wall, as well as the final blended output resulting from
both of our methods.

@@

\appendix
\section{Occlusion Masking}
\label{sec:occlusionMasking}
For the seam minimization process, we need to ensure that our images contain only content
that should be mapped onto the target plane in question. The tiling
approach only checks occlusion for each tile as it is
being textured. For seam minimization, entire images are used, so occlusion checks need to be performed over the entirety of each image to
determine available areas for texture mapping.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, image projections will remain largely rectangular. We can
be efficient by recursively splitting each image into rectangular
pieces, and performing the same occlusion checks used in the tiling
process where needed. To occlude out rectangles, their texture is removed, as untextured areas are
never used for texture mapping.

>>>>

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider a
human-operated backpack system with a number of laser range scanners
as well as 2 cameras facing left and right, each equipped with fisheye
lenses reaching an approximately 180$^{\circ}$ field of view. With the
laser scanners active, the human operator wearing the backpack takes
great care to walk a path such that every wall in the desired indoor
environment is traversed and scanned lengthwise at least once.

Using data collected by the onboard sensors and applying multiple
localization and loop-closure algorithms \cite{chen2010indoor,
  kua2012loopclosure, liu2010indoor}, the backpack is then localized
over its data collection period. This requires recovering the 6
degrees of freedom for the backpack as well as all its sensors,
including the cameras mounted on it\footnote{In this paper, we use the
  terms localization and pose recovery interchangeably, in that they
  both refer to recovering position and orientation.}. Once this is
complete, the data from the laser range scanners is used to generate a
3D point cloud of the surrounding environment.  Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model
\cite{sanchez2012point}. This model, consisting of multiple 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's cameras and their noisy 3D poses, can be considered the
input to our texture mapping problem.

Our problem is different from others in that we are more interested in
creating accurate and seamless textures than in recovering geometry
out of our photos, as is a main concern in many related works
\cite{bernardinimultiplescans, debevechybrid,
  liu2010indoor}. Additionally, our surfaces to be textured are planes
captured from close distances, so we do not always have many visual
reference points to work with, as is the case when texturing detailed objects
at high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or when creating
large scene panoramas \cite{agarwalapanoramas, bernardinimultiplescans,
  debevecviewdependent}. Instead, we need to adjust image projections
relative to each other, while still taking the noisy camera
poses provided by our localization algorithms into account. Using these camera
poses, we also aim for complete automation and improved runtime over
non-localized processes, such as structure from motion or image-based
rendering \cite{agarwalapanoramas, debevechybrid,
  snavelyphototourism}.

@@

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane \cite{rayintersection}.

@@

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{C})
\cdot \vec{N}$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_naive.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache_shift.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_dynprog_noblend.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache_shift_blend.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_dynprog.jpg}}
  \caption{(a): Direct mapping. (b): Mapping with caching. (c):
    Mapping with caching after image alignment. (d): Seam minimization after image alignment). (e): (c) after blending. (f): (d) after blending.}
  \label{fig:compareAll}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while selecting the best image for a given tile. By
using the same image across tile boundaries, we can eliminate a
discontinuity altogether. If this is not possible because a tile is
not visible in images chosen by its neighbors, using similar images is
likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which of these images can map to $t$, and then of
those, we make a choice according to the scoring function in Figure
\ref{fig:scoringFunction}. Before reusing this image, we ensure it
meets a threshold, which we set unconditionally to $\alpha <
45^\circ$, to be considered a viable image, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we then check
our second subset of images, which consists of images that were taken
near the images in the first subset, both spatially and
temporally. These images are not the same as the ones used for
neighboring tiles, but they were taken at a similar location and time,
suggesting that their localization and projection are very
similar. Again, if no good image is found according to the same
threshold, we must then search the entire set of candidate images,
selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

@@

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching (Section
\ref{sec:directMapping}. This set of images constitutes a good
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on our plane is covered by
at least one image in this set; this ensures no holes in our final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring good camera
distances and angles. Third, as a side result of the scoring function,
selected images are only viable candidates for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards
refining and combining their projections.

@@

Figure \ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package which performs homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane, despite the fact that it was
generated after careful hand tuning. Many planes with fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

@@

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we project these images onto the plane
and perform rotation and shifting in order to maximize SIFT feature
matches in overlapping areas between these projections. We then select
which textures to be used either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method to be described in Section \ref{sec:seamMinimization}. This
choice depends on the availability of head-on images for a given
plane. We then finish by applying linear alpha blending to smooth out
any remaining seams.

@@

\subsection{Image Pose Refinement}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images using
SIFT feature matching \cite{lowe1999object}. SIFT matches
determine $dx$ and $dy$ distances between each pair of features for
two images on the plane, though these distances may not always be the
same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that our SIFT-based distances
are reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $dx_{i,j}$ and $dy_{i,j}$ distances
between two images $i$ and $j$. The RANSAC framework handles the
consensus-building machinery, and requires a fitting function and a
distance function. For this application, the fitting function simply
finds the average distance between matches in a pair of images. The
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel outlier
threshold, so that SIFT matches are labeled as outliers if
their horizontal or vertical distances are not within 10 pixels of the
average distance computed by the fitting function.

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches were found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) compare the tile
caching method against this seam minimization method. In Figure
\ref{fig:compareAll}(d), we arbitrarily chose one image for texturing
where images overlap, as blending will be discussed in the next
section. Though both methods provide aligned texturing, the seam
minimization approach results in fewer visible discontinuities, since
it directly reduces the cost of each image boundary, while the tile
caching method uses a scoring function that only approximates this
effect. Furthermore, seam minimization guarantees the best selection
of images, while the sequential tile caching method may select images
early on that turn out to be poor choices once subsequent tiles have
been processed. The seam minimization approach is also far less intensive
in terms of memory usage and runtime, as it does not require
discretizing planes or images.

@@

% \begin{figure*}
%   \centering
%   \includegraphics[width=6in]{pipelineimages.pdf}
%   \caption{Textures corresponding to each step in our texture
%   mapping pipeline.}
%   \label{fig:pipelineimages}
% \end{figure*}

@@

After blending pixels linearly across overlapping regions using alpha
blending, texture mapping is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the final blended output resulting
from both of our methods.

@@

\appendix
\section{Occlusion Masking}
\label{sec:occlusionMasking}
For the seam minimization process, it is important that images contain
only content that should be mapped onto the target plane in
question. The tiling approach checks occlusion for each tile as it is
being textured, whereas the seam minimization approach uses entire
images.  Occlusion checks thus need to be performed over the entirety
of each image to determine available areas for texture mapping.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, image projections will remain largely rectangular. We can be
efficient by recursively splitting each image into rectangular pieces,
and performing the same occlusion checks used in the tiling process
where needed. To occlude out rectangular pieces, their texture is
removed, as untextured areas are never used for texture mapping.
